Swarm coordination through perimeter detection

### 2017-03-22 paper.tex
Fixed References to thesis

This thesis aims to produce algorithms that will reduce the number of active sensors that are needed for the directional coordination of a swarm. When reducing the number of active sensors the algorithms must maintain sufficient coordination of agents to influence the swarm's direction and therefore create an overall \textit{destination vector} to a swarms movement. Reducing the number of active sensors will reduce the net energy usage of the sensors and potentially increase the sensors viable lifetime. Another aspect of the algorithms is that they should minimise the internal disturbance (jitter) created by the \textit{destination vector}.

@@

The following assumptions are made in this thesis:

@@

\section{Swarm destination vector}
The following assumptions are made in this thesis:

@@

In this thesis three algorithms are presented and compared. In the first algorithm, all agents have the role of a coordinator permanently. This will be referred to as the all-agent algorithm. The other two mechanisms employ a selection criteria to identify if the simulated GPS sensor should be enabled. These two algorithms are: full-perimeter detection and the partial perimeter detection. Partial perimeter detection will be referred to as the basic-count algorithm.

There are many other ways in which the coordinator role could be established such as randomly enabling a GPS for a set period of time or oscillating the role on and off but they are beyond the scope of this thesis.

@@

This problem can be resolved fully by introducing a communications channel that would allow the affected agents to negotiate a resolution to chose one of the shorter pathways~\cite{MD:09}, as there is no communications available this is not an option. In this thesis the error is limited to a very specific set of circumstances which will only occur during the initial expansion phase or during obstacle avoidance. The algorithm described in~Section~\ref{sec:PerimeterAgentDetection} selects both agents as being perimeter agents therefore the problem can is resolved by using the pseudo-perimeter ($5 \rightarrow b^{'} \rightarrow b^{''} \rightarrow 1$) ensuring a continuous perimeter is detected.

@@

\subsubsection{Perimeter and void detection}\label{sec:PerimeterVoidDetection} 
The focus of this thesis is the control of arbitrary sized swarms. The identification of the inner and outer perimeters would require a communications mechanism to determine which type of perimeter an agent was a part of~\cite{ZAPS:07,MD:09}. A mean angle $< 180^\circ$ would indicate a void. A mean angle $> 180^\circ$ would indicate an outer perimeter which may be enclose inside a swarm as shown in~Figure~\ref{concave:VoidPerimeter2} on page \pageref{concave:VoidPerimeter2}.

@@

A further issue of introducing a communications infrastructure would be the increase in the energy usage of the swarm. Identification of the perimeter type is therefore not practical in context of this thesis.

@@

The perimeter detection algorithms employed in this thesis have been reduced from ($O(n^2)$) to zero by limiting the agents information needs to simple proximity information from a sensor. This is a bio-inspired approach taken from the flocking of birds and shoaling of fish.

By reducing the message propagation to zero the scalability of the perimeter and void reduction algorithms are improved. Most research using message propagation techniques use swarms of at most 50 agents, as discussed this thesis is based upon the control of arbitrary sized swarms.

>>>>

This paper aims to produce algorithms that will reduce the number of active sensors that are needed for the directional coordination of a swarm. When reducing the number of active sensors the algorithms must maintain sufficient coordination of agents to influence the swarm's direction and therefore create an overall \textit{destination vector} to a swarms movement. Reducing the number of active sensors will reduce the net energy usage of the sensors and potentially increase the sensors viable lifetime. Another aspect of the algorithms is that they should minimise the internal disturbance (jitter) created by the \textit{destination vector}.

@@

The following assumptions are made in this paper:

@@

\section{Swarm destination vector}
The following assumptions are made in this paper:

@@

In this paper three algorithms are presented and compared. In the first algorithm, all agents have the role of a coordinator permanently. This will be referred to as the all-agent algorithm. The other two mechanisms employ a selection criteria to identify if the simulated GPS sensor should be enabled. These two algorithms are: full-perimeter detection and the partial perimeter detection. Partial perimeter detection will be referred to as the basic-count algorithm.

There are many other ways in which the coordinator role could be established such as randomly enabling a GPS for a set period of time or oscillating the role on and off but they are beyond the scope of this paper.

@@

This problem can be resolved fully by introducing a communications channel that would allow the affected agents to negotiate a resolution to chose one of the shorter pathways~\cite{MD:09}, as there is no communications available this is not an option. In this paper the error is limited to a very specific set of circumstances which will only occur during the initial expansion phase or during obstacle avoidance. The algorithm described in~Section~\ref{sec:PerimeterAgentDetection} selects both agents as being perimeter agents therefore the problem can is resolved by using the pseudo-perimeter ($5 \rightarrow b^{'} \rightarrow b^{''} \rightarrow 1$) ensuring a continuous perimeter is detected.

@@

\subsubsection{Perimeter and void detection}\label{sec:PerimeterVoidDetection} 
The focus of this paper is the control of arbitrary sized swarms. The identification of the inner and outer perimeters would require a communications mechanism to determine which type of perimeter an agent was a part of~\cite{ZAPS:07,MD:09}. A mean angle $< 180^\circ$ would indicate a void. A mean angle $> 180^\circ$ would indicate an outer perimeter which may be enclose inside a swarm as shown in~Figure~\ref{concave:VoidPerimeter2} on page \pageref{concave:VoidPerimeter2}.

@@

A further issue of introducing a communications infrastructure would be the increase in the energy usage of the swarm. Identification of the perimeter type is therefore not practical in context of this paper.

@@

The perimeter detection algorithms employed in this paper have been reduced from ($O(n^2)$) to zero by limiting the agents information needs to simple proximity information from a sensor. This is a bio-inspired approach taken from the flocking of birds and shoaling of fish.

By reducing the message propagation to zero the scalability of the perimeter and void reduction algorithms are improved. Most research using message propagation techniques use swarms of at most 50 agents, as discussed this paper is based upon the control of arbitrary sized swarms.
